### 2014-09-27 rebuttal.md

3. The author mentioned in results line 185 “The kidney appears to [be] an outlier in the number of unique sequences, though this could […] result [from] the recovery of more lowly expressed transcripts [caused by] deeper sequencing.” Why would this not also be the case for liver, which only has 3M (5%) less sequences?

	*I have removed this statement. I do wonder why this organ produced a larger number of contigs in the assembly, but a similar number of transcripts after filtering. I admit that I do not have a good explanation, and I hope a reader will..*

4. I am trying to understand the filtering process for the assembled reads. From my understanding (Page 4, lines 103:109) sequences were filtered using Blastn, (Page 4, lines 113:120) annotated using Blastn, HMMER3 and Transdecoder. Is my understanding correct? If so, why were the assembled sequences filtered with Blastn before annotated with Blastn and HMMER3? I thought the point of HMMER3 was to retain divergent sequences not detected by blastn.

	*Yes, I filtered based on blastN to the Peromyscus maniculatus and Mus musculus transcriptomes, as well as a Mus ncRNA dataset. Given the combo of P. maniculatus being so closely related (and with a high quality annotation) and Mus being more distantly related but fantastically annotated coding regions, I am confident that I am not missing a substantial number of 'real' transcripts by employing this strategy. I think it is likely that I could have recovered more contigs using Pfam, but I would worry that the sensitivity of a HMM based search may result in the recovery of 'false' transcripts, which I am relatively intolerant of. I am electing to not filter based on Pfam, understanding the implications of this.*

5. For the natural section results, I think it would be interest to add more than two genes. Perhaps the top and bottom 10 genes from the Tajima’s D analysis.

	*I have added a few more genes, Aqp1,2,4,9. There is not evidence of positive selection in the eremicus lineage for any of these genes, as expected. The de novo (e.g. NOT focused on candidate genes) exploration of positive selection on P. eremicus will be presented in a later work, in connection with a genome sequence*

6. It would also be nice to have the various parts of the analysis in a repository, for reviewing and open science purposes.

	*I will make sure that the scripts are placed in a public GitHub repo.*

### 2014-09-27 Pero_Transcriptome.tex
changed ref. style

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
\documentclass[11pt]{article}
%\usepackage{natbib}
\usepackage{hyperref}
\usepackage{lineno}

@@

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

@@

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

@@

Deserts are widely considered one of the harshest environments on Earth. Animals living in desert environments are forced to endure intense heat and drought, and as a result, species living in these environments are likely to possess specialized mechanisms to deal with them. While living in deserts likely involves a large number of adaptive traits, the ability to osmoregulate -- to maintain the proper water and electrolyte balance -- appears to be paramount \cite{Walsberg:2000uu}. Indeed, the maintenance of water balance is one of the most important physiologic processes for all organisms, whether they be desert inhabitants or not. Most animals are exquisitely sensitive to changes in osmolality, with slight derangement eliciting physiologic compromise.  When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur. Thus there has likely been strong selection for mechanisms supporting optimal osmoregulation in species that live where water is limited. Understanding these mechanisms will significantly enhance our understanding of the physiologic processes underlying osmoregulation in extreme environments, which will have implications for studies of human health, conservation, and climate change. \\

The genes and structures responsible for the maintenance of water and electrolyte balance are well characterized in model organisms such as mice \cite{Tatum:2009kp}, rats \cite{Romero:2007fs,Rojek:2006dg,NIELSEN:1995uq}, and humans \cite{Mobasheri:2007tt,Bedford:2003wa,NIELSEN:1999vg}. These studies, many of which have been enabled by newer sequencing technologies, provide a foundation for studies of renal genomics in non-model organisms. Because researchers have long been interested in desert adaptation, a number of studies have looked at the morphology or expression of single genes in the renal tissues of desert adapted rodents \textit{Phyllotis darwini} \cite{Gallardo:2005fm}, \textit{Psammomys obesus} \cite{Kaissling:1975uu}, and \textit{Perognathus penicillatus} \cite{Altschuler:1979du}. More recently, full renal transcriptomes have been generated for \textit{Dipodomys spectabilis} and \textit{Chaetodipus baileyi}, \cite{Marra:2014de} as well as \textit{Abrothrix olivacea} \cite{Giorello:2014iv}. \\

These studies provide a rich context for current and future work aimed at developing a synthetic understanding of the genetic and genomic underpinnings of desert adaptation in rodents. As a first step, we have sequenced, assembled, and characterized the transcriptome (using four tissue types - liver, kidney, testes and brain) of a desert adapted cricetid rodent endemic to the southwest United States, \textit{Peromyscus eremicus}. These animals have a lifespan typical of small mammals \cite{Veal:2001vp}, and therefore an individual may live its entire life without ever drinking water. Additionally, they have a distinct advantage over other desert animals (e.g. \textit{Dipodomys}) in that they breed readily in captivity, which enables future laboratory studies of the phenotype of interest. In addition, the focal species is positioned in a clade of well known animals (e.g. \textit{P. californicus}, \textit{P. maniculatus}, and \textit{P. polionotus}) \cite{Bradley:2007jo} with growing genetic and genomic resources \cite{Shorter:2014kw,Panhuis:2011co,Shorter:2012ib}. Together, this suggests that future comparative studies are possible. \\

@@

To begin to understand how genes may underlie desert adaptation, we collected 16 adult individuals (9 male, 7 female) from a single population of \textit{P. eremicus} over a two-year time period (2012-2013). These individuals were captured in live traps and then euthanized using isoflurane overdose and decapitation. Immediately post-mortem, the abdominal and pelvic organs were removed, cut in half (in the case of the kidneys), placed in RNAlater and flash frozen in liquid nitrogen. Removal of the brain, with similar preservation techniques, followed. Time from euthanasia to removal of all organs never exceeded five minutes. Samples were transferred to a -80C freezer at a later date. These procedures were approved by the Animal Care and Use Committee located at the University of California Berkeley (protocol number R224) and University of New Hampshire (protocol number 130902) as well as the California Department of Fish and Game (protocol SC-008135) and followed guidelines established by the American Society of Mammalogy for the use of wild animals in research \cite{Sikes:2011dz}.

@@

The raw sequence reads corresponding to the four tissue types were error corrected using the software bless version 0.17 \cite{Heo:2014cb} using kmer=25, based on the developer's default recommendations. The error-corrected sequence reads were adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}. Specifically, adapter sequence contamination and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}. Reads from each tissue were assembled using the Trinity version released 17 July 2014 \cite{Haas:2013jq}. We used flags to indicate the stranded nature of sequencing reads and set the maximum allowable physical distance between read pairs to 999nt. We elected to assembly reads derived from a single deeply sequenced individual (Peer360, a male) to reduce polymorphism and thus the complexity of the de Bruijn graph, which has important implications for runtime, hardware requirements \cite{Lowe:2014iv,Pop:2009dp}, and assembly contiguity \cite{Vijay:2012gy and MacManes unpublished observations}. Individual tissues were assembled independently, as we hypothesize that tissue specific isoforms would be reconstructed with higher fidelity that if all tissues were assembled together.

The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM. To filter the raw sequence assembly, we downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}) and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}). We used a blastN (version 2.2.29+) procedure (default settings, evalue set to 10$^{-10}$) to identify contigs in the \textit{P. eremicus} dataset likely to be biological in origin. This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if transcripts per million (TPM) \textless 1  \cite{MacManes:2012bu}) of the raw assembly. We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \cite{Li:2006hr} using default settings, except that sequences were clustered based on 95\% sequence similarity. This multi-fasta file was used for all subsequent analyses, including annotation and mapping. \\

@@

The filtered assemblies were annotated using the default settings of the blastN algorithm \cite{Camacho:2009fc} against the Ensembl cDNA and ncRNA datasets described above, downloaded on 1 August 2014. Among other things, the Ensemble transcript identifiers were used in the analysis of gene ontology conducted in the PANTHER package \cite{Mi:2004iv}. Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 version 3.1b1 \cite{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \cite{Punta:2012ko}. Lastly, we extracted putative coding sequences using Transdecoder version 4Jul2014 (\url{http://transdecoder.sourceforge.net/})\\

To identify patterns of gene expression unique to each tissue type, we mapped sequence reads from each tissue type to the reference assembly using bwa-mem (version cloned from Github 7/1/2014)  \cite{Li:2013wn}. We estimated expression for the four tissues individually using default settings of the software eXpress version 1.51 \cite{Roberts:2012dh}. Interesting patterns of expression, including instances where expression was limited to a single tissue type, were identified and visualized. \\

@@

In addition to the reference individual sequenced at four different tissue types, we sequenced 15 other conspecific individuals from the same population in Palm Desert, California. Sequence data were mapped to the reference transcriptome using bwa-mem. The alignments were sorted and converted to BAM format, then passed to the program \textsc{Angsd} version 0.610, which was used for calculating the folded site frequency spectrum (SFS) and Tajima's D \cite{Korneliussen:2013uz}. \\

@@

To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a recent work on desert-adapted \textit{Dipodomys} rodents \cite{Marra:2014de}. The coding sequences corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE version 1.01b\cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} as identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}). An unrooted gene tree with branch lengths was constructed using the online resource ClustalW2-Phylogeny (\url{http://www.ebi.ac.uk/Tools/phylogeny/clustalw2_phylogeny/}), and the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \cite{Yang:2011bm,Yang:2007ki}. Significance was evaluated via the use of the likelihood ratio test. \\

@@

Transcriptome assemblies for each tissue type was accomplished using the program Trinity \cite{Haas:2013jq}. The raw assemblies for brain, liver, testes, and kidney contained 185425, 222096, 180233, and 514091 assembled sequences respectively. This assembly was filtered using a blastN procedure against the \textit{Mus} cDNA and ncRNA and \textit{P. maniculatus} cDNAs, which resulted in a final dataset containing 68331 brain-derived transcripts, 71041 liver-derived transcripts, 67340 testes-derived transcripts, and 113050 kidney-derived transcripts. Mapping the error-corrected adapter/quality trimmed reads to these datasets resulted in mapping 94.98\% (87.01\% properly paired) of the brain-derived reads to the brain transcriptome, 96.07\% (88.13\% properly paired) of the liver-derived reads to the liver transcriptome, 96.81\% (85.10\% properly paired) of the testes-derived reads to the testes transcriptome, and 91.87\% (83.77\% properly paired) of the kidney-derived reads to the kidney transcriptome. Together, these statistics suggest that the tissue-specific transcriptomes are of extremely high quality. All tissue-specific assemblies are to be made available on Dryad, and until then are stored on Dropbox (\url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0}). \\

@@

The distribution of the estimates of Tajima's D for all of the assembled transcripts is shown in Figure 2. The distribution is skewed toward negative values (mean=-0.89, variance=0.58), which is likely the result of purifying selection, a model of evolution commonly invoked for coding DNA sequences \cite{Chamary:2006db}. \hyperlink{Table 3}{Table 3} presents the 10 transcripts whose estimate of Tajima's D is the greatest, while \hyperlink{Table 4}{Table 4} presents the 10 transcripts whose estimate of Tajima's D is the least. The former list of genes is likely to contain transcripts experiencing balancing selection in the studied population. This list includes, interestingly, genes obviously related to solute and water balance (e.g. Clcnkb and a transmembrane protein gene) and immune function (a interferon-inducible GTPase and a Class 1 MHC gene). The latter group, containing transcripts whose estimates of Tajima's D are the smallest are likely experiencing purifying selection. Many of these transcripts are involved in core regulatory functions where mutation may have strongly negative fitness consequences. \\

@@

To begin to test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, we implemented the branch-site test as described above using alignments produced in MACSE. These alignments were free from indels and internal stop codons. We set the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions. These two transcripts were chosen specifically because they - the former significantly - were recently linked to osmoregulation in a desert rodent \cite{Marra:2014de}. The test for Slc2a9 was highly significant (2$\Delta$Lnl=51.4, df=1, p=0), indicating enhanced selection in \peer\: relative to the other lineages. The branch site test for positive selection conducted on the Vdr gene was non-significant (2$\Delta$Lnl=0.68, df=1, p=1). This limited analysis of selection is to be followed up by an analysis of genome wide patterns of natural selection. \\

>>>>

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
\documentclass[12pt]{article}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{lineno}

@@

% cite package, to clean up citations in the main text. Do not remove.
%\usepackage{cite}

@@

% Use the PLoS provided bibtex style
\bibliographystyle{apalike}

@@

Deserts are widely considered one of the harshest environments on Earth. Animals living in desert environments are forced to endure intense heat and drought, and as a result, species living in these environments are likely to possess specialized mechanisms to deal with them. While living in deserts likely involves a large number of adaptive traits, the ability to osmoregulate -- to maintain the proper water and electrolyte balance -- appears to be paramount \citep{Walsberg:2000uu}. Indeed, the maintenance of water balance is one of the most important physiologic processes for all organisms, whether they be desert inhabitants or not. Most animals are exquisitely sensitive to changes in osmolality, with slight derangement eliciting physiologic compromise.  When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur. Thus there has likely been strong selection for mechanisms supporting optimal osmoregulation in species that live where water is limited. Understanding these mechanisms will significantly enhance our understanding of the physiologic processes underlying osmoregulation in extreme environments, which will have implications for studies of human health, conservation, and climate change. \\

The genes and structures responsible for the maintenance of water and electrolyte balance are well characterized in model organisms such as mice \citep{Tatum:2009kp}, rats \citep{Romero:2007fs,Rojek:2006dg,NIELSEN:1995uq}, and humans \citep{Mobasheri:2007tt,Bedford:2003wa,NIELSEN:1999vg}. These studies, many of which have been enabled by newer sequencing technologies, provide a foundation for studies of renal genomics in non-model organisms. Because researchers have long been interested in desert adaptation, a number of studies have looked at the morphology or expression of single genes in the renal tissues of desert adapted rodents \textit{Phyllotis darwini} \citep{Gallardo:2005fm}, \textit{Psammomys obesus} \citep{Kaissling:1975uu}, and \textit{Perognathus penicillatus} \citep{Altschuler:1979du}. More recently, full renal transcriptomes have been generated for \textit{Dipodomys spectabilis} and \textit{Chaetodipus baileyi}, \citep{Marra:2014de} as well as \textit{Abrothrix olivacea} \citep{Giorello:2014iv}. \\

These studies provide a rich context for current and future work aimed at developing a synthetic understanding of the genetic and genomic underpinnings of desert adaptation in rodents. As a first step, we have sequenced, assembled, and characterized the transcriptome (using four tissue types - liver, kidney, testes and brain) of a desert adapted cricetid rodent endemic to the southwest United States, \textit{Peromyscus eremicus}. These animals have a lifespan typical of small mammals \citep{Veal:2001vp}, and therefore an individual may live its entire life without ever drinking water. Additionally, they have a distinct advantage over other desert animals (e.g. \textit{Dipodomys}) in that they breed readily in captivity, which enables future laboratory studies of the phenotype of interest. In addition, the focal species is positioned in a clade of well known animals (e.g. \textit{P. californicus}, \textit{P. maniculatus}, and \textit{P. polionotus}) \citep{Bradley:2007jo} with growing genetic and genomic resources \citep{Shorter:2014kw,Panhuis:2011co,Shorter:2012ib}. Together, this suggests that future comparative studies are possible. \\

@@

To begin to understand how genes may underlie desert adaptation, we collected 16 adult individuals (9 male, 7 female) from a single population of \textit{P. eremicus} over a two-year time period (2012-2013). These individuals were captured in live traps and then euthanized using isoflurane overdose and decapitation. Immediately post-mortem, the abdominal and pelvic organs were removed, cut in half (in the case of the kidneys), placed in RNAlater and flash frozen in liquid nitrogen. Removal of the brain, with similar preservation techniques, followed. Time from euthanasia to removal of all organs never exceeded five minutes. Samples were transferred to a -80C freezer at a later date. These procedures were approved by the Animal Care and Use Committee located at the University of California Berkeley (protocol number R224) and University of New Hampshire (protocol number 130902) as well as the California Department of Fish and Game (protocol SC-008135) and followed guidelines established by the American Society of Mammalogy for the use of wild animals in research \citep{Sikes:2011dz}.

@@

The raw sequence reads corresponding to the four tissue types were error corrected using the software bless version 0.17 \citep{Heo:2014cb} using kmer=25, based on the developer's default recommendations. The error-corrected sequence reads were adapter and quality trimmed following recommendations from MacManes \citep{MacManes:2014io} and Mbandi \citep{Christoffels:2014bg}. Specifically, adapter sequence contamination and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \citep{Lohse:2012fg}. Reads from each tissue were assembled using the Trinity version released 17 July 2014 \citep{Haas:2013jq}. We used flags to indicate the stranded nature of sequencing reads and set the maximum allowable physical distance between read pairs to 999nt. We elected to assembly reads derived from a single deeply sequenced individual (Peer360, a male) to reduce polymorphism and thus the complexity of the de Bruijn graph, which has important implications for runtime, hardware requirements \citep{Lowe:2014iv,Pop:2009dp}, and assembly contiguity \citep{Vijay:2012gy and MacManes unpublished observations}. Individual tissues were assembled independently, as we hypothesize that tissue specific isoforms would be reconstructed with higher fidelity that if all tissues were assembled together.

The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM. To filter the raw sequence assembly, we downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}) and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}). We used a blastN (version 2.2.29+) procedure (default settings, evalue set to 10$^{-10}$) to identify contigs in the \textit{P. eremicus} dataset likely to be biological in origin. This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if transcripts per million (TPM) \textless 1  \citep{MacManes:2012bu}) of the raw assembly. We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \citep{Li:2006hr} using default settings, except that sequences were clustered based on 95\% sequence similarity. This multi-fasta file was used for all subsequent analyses, including annotation and mapping. \\

@@

The filtered assemblies were annotated using the default settings of the blastN algorithm \citep{Camacho:2009fc} against the Ensembl cDNA and ncRNA datasets described above, downloaded on 1 August 2014. Among other things, the Ensemble transcript identifiers were used in the analysis of gene ontology conducted in the PANTHER package \citep{Mi:2004iv}. Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 version 3.1b1 \citep{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \citep{Punta:2012ko}. Lastly, we extracted putative coding sequences using Transdecoder version 4Jul2014 (\url{http://transdecoder.sourceforge.net/})\\

To identify patterns of gene expression unique to each tissue type, we mapped sequence reads from each tissue type to the reference assembly using bwa-mem (version cloned from Github 7/1/2014)  \citep{Li:2013wn}. We estimated expression for the four tissues individually using default settings of the software eXpress version 1.51 \citep{Roberts:2012dh}. Interesting patterns of expression, including instances where expression was limited to a single tissue type, were identified and visualized. \\

@@

In addition to the reference individual sequenced at four different tissue types, we sequenced 15 other conspecific individuals from the same population in Palm Desert, California. Sequence data were mapped to the reference transcriptome using bwa-mem. The alignments were sorted and converted to BAM format, then passed to the program \textsc{Angsd} version 0.610, which was used for calculating the folded site frequency spectrum (SFS) and Tajima's D \citep{Korneliussen:2013uz}. \\

@@

To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a recent work on desert-adapted \textit{Dipodomys} rodents \citep{Marra:2014de}. The coding sequences corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE version 1.01b \citep{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} as identified by the conditional reciprocal best blast procedure (CRBB, \citep{Aubry:2014en}). An unrooted gene tree with branch lengths was constructed using the online resource ClustalW2-Phylogeny (\url{http://www.ebi.ac.uk/Tools/phylogeny/clustalw2_phylogeny/}), and the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \citep{Yang:2011bm,Yang:2007ki}. Significance was evaluated via the use of the likelihood ratio test. \\

@@

Transcriptome assemblies for each tissue type was accomplished using the program Trinity \citep{Haas:2013jq}. The raw assemblies for brain, liver, testes, and kidney contained 185425, 222096, 180233, and 514091 assembled sequences respectively. This assembly was filtered using a blastN procedure against the \textit{Mus} cDNA and ncRNA and \textit{P. maniculatus} cDNAs, which resulted in a final dataset containing 68331 brain-derived transcripts, 71041 liver-derived transcripts, 67340 testes-derived transcripts, and 113050 kidney-derived transcripts. Mapping the error-corrected adapter/quality trimmed reads to these datasets resulted in mapping 94.98\% (87.01\% properly paired) of the brain-derived reads to the brain transcriptome, 96.07\% (88.13\% properly paired) of the liver-derived reads to the liver transcriptome, 96.81\% (85.10\% properly paired) of the testes-derived reads to the testes transcriptome, and 91.87\% (83.77\% properly paired) of the kidney-derived reads to the kidney transcriptome. Together, these statistics suggest that the tissue-specific transcriptomes are of extremely high quality. All tissue-specific assemblies are to be made available on Dryad, and until then are stored on Dropbox (\url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0}). \\

@@

The distribution of the estimates of Tajima's D for all of the assembled transcripts is shown in Figure 2. The distribution is skewed toward negative values (mean=-0.89, variance=0.58), which is likely the result of purifying selection, a model of evolution commonly invoked for coding DNA sequences \citep{Chamary:2006db}. \hyperlink{Table 3}{Table 3} presents the 10 transcripts whose estimate of Tajima's D is the greatest, while \hyperlink{Table 4}{Table 4} presents the 10 transcripts whose estimate of Tajima's D is the least. The former list of genes is likely to contain transcripts experiencing balancing selection in the studied population. This list includes, interestingly, genes obviously related to solute and water balance (e.g. Clcnkb and a transmembrane protein gene) and immune function (a interferon-inducible GTPase and a Class 1 MHC gene). The latter group, containing transcripts whose estimates of Tajima's D are the smallest are likely experiencing purifying selection. Many of these transcripts are involved in core regulatory functions where mutation may have strongly negative fitness consequences. \\

@@

To begin to test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, we implemented the branch-site test as described above using alignments produced in MACSE. These alignments were free from indels and internal stop codons. We set the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions. These two transcripts were chosen specifically because they - the former significantly - were recently linked to osmoregulation in a desert rodent \citep{Marra:2014de}. The test for Slc2a9 was highly significant (2$\Delta$Lnl=51.4, df=1, p=0), indicating enhanced selection in \peer\: relative to the other lineages. The branch site test for positive selection conducted on the Vdr gene was non-significant (2$\Delta$Lnl=0.68, df=1, p=1). This limited analysis of selection is to be followed up by an analysis of genome wide patterns of natural selection. \\
